---
title: "Missclasification matrix"
author: "Emilio Otero"
date: "2026-03-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("stat_func.R")
source("pupulation_model.R")
source("thikanov_inv.R")
source("simulation_data.R")
library(xtable)
library(glmnet)
library(purrr)
library(patchwork)
library(ggplot2)
library(dplyr)
library(knitr)
library(expm)
library(here)


```

```{r}
# My missclass matrix
missclass_matrix_A <-  truncated_misclassification(func_sigma = function(x) 1)
missclass_matrix_B <-  truncated_misclassification(func_sigma = function(x) 4)
missclass_matrix_C <-  truncated_misclassification(func_sigma = function(x) 1, func_mu = function(x) x+4/(x+0.5))
missclass_matrix_D <-  truncated_misclassification(func_sigma = function(x) 4, func_mu = function(x) x+4/(x+0.5))

missclass_all_t <- list(
  A = t(missclass_matrix_A),
  B = t(missclass_matrix_B),
  C = t(missclass_matrix_C),
  D = t(missclass_matrix_D)
)

missclass_all <- list(
  A = missclass_matrix_A,
  B = missclass_matrix_B,
  C = missclass_matrix_C,
  D = missclass_matrix_D
)

# Set samples
samples <- c(10, 25, 50, 100, 150, 200)
```



## Plot stable age distribution

```{r}
# Ensure folder exists
dir.create("figures", showWarnings = FALSE)


# Create plot
p1 <- ggplot(dist_eigenvector, aes(x = age_labels, y = stable_dist)) +  # change 'value' if needed
  geom_col() +
  labs(
    x = "Age class",
    y = "Probability",
    title = "Stable age distribution"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)  # rotate 45 degrees
  )

# Save automatically
ggsave(
  filename = "../figures/age_distribution_p1.pdf",
  plot = p1,
  width = 6,
  height = 4
)

p1
```


## naiv metod (mätfel)
```{r}


# convolute distribution analytically  with a msicclass matrix
conv_data <- as.vector(t(missclass_matrix_D)%*%dist_eigenvector$stable_dist)

dist1 <- data.frame(
  age_labels = dist_eigenvector$age_labels,
  stable_dist = dist_eigenvector$stable_dist
) %>%
  mutate(source = "Stable age distribution")

dist2 <- data.frame(
  age_labels = dist_eigenvector$age_labels,  # same x-axis!
  stable_dist = conv_data
) %>%
  mutate(source = "Convoluted distribution")

combined <- bind_rows(dist2, dist1)

p2 <- ggplot(combined, aes(x = age_labels, y = stable_dist, fill = source)) +
  geom_col(position = "dodge", alpha = 0.5) +  
  scale_fill_manual(
    name = "Distribution type",   # 👈 legend title
    values = c(
      "Convoluted distribution" = "Gray40",
      "Stable age distribution" = "Black"
    ) 
  ) +
  labs(
    x = "Age class",
    y = "Probability",
    title = "Stable age vs convoluted age distribution"
  ) +
  theme_minimal() +
    theme(
    legend.position = c(0.8, 0.8),  # x=0.8 (right), y=0.8 (top)
    legend.background = element_rect(fill = "white", color = "black", size = 0.3),
    legend.key.size = unit(0.6, "lines")
  )


# Save automatically
ggsave(
  filename = "../figures/age_distribution_comp_p2.pdf",
  plot = p2,
  width = 6,
  height = 4
)
p2

```
## Numerisk instabilitet plot

```{r}

# Standard deviations from 0.2 to 20
sigma_values <- seq(0.2, 10, length.out = 1000)

  

# Function to calculate condition number for a given sigma
calc_condition <- function(sigma) {
  mat <- truncated_misclassification(
    func_sigma = function(x) sigma,
    func_mu    = function(x) x
  )
  # Compute condition number
  kappa(mat)
}

# Apply function over all sigma values
cond_numbers <- map_dbl(sigma_values, calc_condition)

plot_data <- data.frame(
  sigma = sigma_values,
  variance = sigma_values^2,
  condition_number = cond_numbers
)

thresholds <- round(log(c(1, 5, 100)),3)

p3 <- ggplot(plot_data, aes(x = sigma, y = condition_number)) +
  geom_point(color = "black", size = 0.5) +
  geom_hline(yintercept = thresholds, linetype = "dashed", color = "gray40") +
  annotate("text", x = c(0.5,2,3.5), y = thresholds, label = paste0("κ=", c(1,5,100)),
           hjust = 0, vjust = -0.3, size = 3.5, color = "gray40") +
  scale_y_log10() +
  labs(
    x = "Standard Deviation (σ)",
    y = "Logged Condition Number (κ)",
    title = "Logged Condition Number vs standard deviation of error"
  ) +
  theme_minimal()


ggsave(
  filename = "../figures/kondition_number_vs_sd_p3.pdf",
  plot = p3,
  width = 6,
  height = 4
)
p3
```

## True distribution sampling & statistics & table
```{r}
# simulate 

sim_results_true_dist <- lapply( list(stable_dist), function(dist) {
  lapply(samples, function(n) {
    simulate_counts(dist, size_per_sample = n, n_samples = 100)
  })
})

#  summary statistic of TV from sim_results
tv_result_summary_true <-  compute_tv_all(sim_results_true_dist, stable_dist)

# convert to data frame
df_tv_true <- tv_to_df_summary(tv_result_summary_true)
df_tv_true$model <- "True distribution"
df_tv_true

# Export to ..\tables
table_tex <- kable(df_tv_true, format = "latex", booktabs = TRUE)

writeLines(table_tex, "../tables/df_tv_true_t1.tex")
```


### Convolute samples 
```{r}
# Show convolution 
set.seed(19911122)

# Convoluted Distribution
conv_dist <- lapply(missclass_all_t, function(M) {
  as.vector(M %*% stable_dist)
})


# To each type of conv matrix, sample for each type of sample size 
# Dist A,B,C,D = 1,2,3,4 & sampel size 10, 25, 50, 100 = 1, 2, 3, 4,5,6 Each one is a matrix where column is replica and row "bin" 
sim_results_c <- lapply(conv_dist, function(dist) {
  lapply(samples, function(n) {
    simulate_counts(dist, size_per_sample = n, n_samples = 100)
  })
})

#  summary statistic of TV from sim_results
tv_result_summary <-  compute_tv_all(convolved_sim, stable_dist)

# convert to data frame
df_tv_c <- tv_to_df_summary(tv_result_summary)
df_tv_c


# Export to ..\tables
table_tex <- kable(df_tv_c, format = "latex", booktabs = TRUE)

writeLines(table_tex, "../tables/df_tv_convolution_t2.tex")

```

### Convolution graph
```{r}

df_all_c_t <- rbind(df_tv_c, df_tv_true)

p4 <- ggplot(df_all_c_t, aes(x = sample_size, y = mean, group = model, color = model)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(
    title = "Convergence of TV Distance vs sample size",
    x = "Sample size",
    y = "Mean TV distance",
    color = "Error Model"
  ) +
  scale_color_manual(values = c(
    "A" = "red",
    "B" = "blue",
    "C" = "darkgreen",
    "D" = "orange",
    "True distribution" = "black"
  )) +
  coord_cartesian(ylim = c(0, 0.7)) +
  theme_minimal(base_size = 14)



ggsave(
  filename = "../figures/Conv_TV_vs_sample_p4.pdf",
  plot = p4,
  width = 6,
  height = 4
)



p4
```
### Deconvolution uniform-init simulation 
```{r}
# Deconvolution steps
set.seed(19911122)

deconvolved_samples_u <- Map(function(dist_list, K) {
  lapply(dist_list, function(mat) {
    apply(mat, 2, function(col) {
      em_deconvolve(col, K = K)$p
    })
  })
}, sim_results_c, missclass_all)

#  summary statistic of TV from sim_results
tv_result_summary_deconv <-  compute_tv_all(deconvolved_samples_u, stable_dist)

# convert to data frame
df_tv_deconv_u <- tv_to_df_summary(tv_result_summary_deconv)

df_tv_deconv_u

# Export to ..\tables
table_tex <- kable(df_tv_deconv_u, format = "latex", booktabs = TRUE)

writeLines(table_tex, "../tables/df_tv_deconvolution_u_t3.tex")
# 
# A	10	0.5635416	0.5442254	0.7641923
# A	25	0.4792911	0.4671815	0.7003950
# A	50	0.3984489	0.3892980	0.5646527
# A	100	0.3354296	0.3257366	0.4896924
# A	150	0.2741892	0.2598381	0.3893600
# A	200	0.2690260	0.2600026	0.3836192
# B	10	0.5363998	0.5455556	0.7712813
# B	25	0.4087443	0.4013897	0.6919004
# B	50	0.3180364	0.3045640	0.5465116
# B	100	0.2502946	0.2323665	0.4132018
```

### Deconvolution uniform graph
```{r}

df_all_du_t <- rbind(df_tv_deconv_u, df_tv_true)

p5 <- ggplot(df_all_du_t, aes(x = sample_size, y = mean, group = model, color = model)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(

    x = "Sample size",
    y = "Mean TV distance",
    color = "Error Model"
  ) +
  scale_color_manual(values = c(
    "A" = "red",
    "B" = "blue",
    "C" = "darkgreen",
    "D" = "orange",
    "True distribution" = "black"
  )) +
  coord_cartesian(ylim = c(0, 0.7)) +
  theme_minimal(base_size = 14)

p5

ggsave(
  filename = "../figures/deconv_TV_vs_sample_u_p5.pdf",
  plot = p5,
  width = 6,
  height = 4
)

```
## Deconvolution well-informed init simulation

```{r}
# Deconvolution steps
set.seed(19911122)

deconvolved_samples_i <- Map(function(dist_list, K) {
  lapply(dist_list, function(mat) {
    apply(mat, 2, function(col) {
      em_deconvolve(col, K = K, init = TRUE)$p
    })
  })
}, sim_results_c, missclass_all)

#  summary statistic of TV from sim_results
tv_result_summary_deconv_i <-  compute_tv_all(deconvolved_samples_i, stable_dist)

# convert to data frame
df_tv_deconv_i <- tv_to_df_summary(tv_result_summary_deconv_i)

df_tv_deconv_i

# Export to ..\tables
table_tex <- kable(df_tv_deconv_i, format = "latex", booktabs = TRUE)

writeLines(table_tex, "../tables/df_tv_deconvolution_i_t3.tex")
```
### Deconvolution informed - Graph
```{r}
df_all_di_i <- rbind(df_tv_deconv_i, df_tv_true)

p6 <- ggplot(df_all_di_i, aes(x = sample_size, y = mean, group = model, color = model)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(
    x = "Sample size",
    y = "Mean TV distance",
    color = "Error Model"
  ) +
  scale_color_manual(values = c(
    "A" = "red",
    "B" = "blue",
    "C" = "darkgreen",
    "D" = "orange",
    "True distribution" = "black"
  )) +
  coord_cartesian(ylim = c(0, 0.7)) +
  theme_minimal(base_size = 14)


p6

ggsave(
  filename = "../figures/deconv_TV_vs_sample_i_p6.pdf",
  plot = p6,
  width = 6,
  height = 4
)
  
```

```{r}
p7 <- (p5 | p6) +
  plot_layout(guides = "collect")  + 
  plot_annotation(
    title = "Convergence of TV Distance vs sample size",
    tag_levels = "a"
  ) &
  theme(legend.position = "bottom")

p7

ggsave(
  filename = "../figures/deconv_TV_vs_sample_p7.pdf",
  plot = p7,
  width = 6,
  height = 4
)
```

```{r}

df_diff_i <- df_tv_deconv_i %>%
  left_join(df_tv_true, by = "sample_size", suffix = c("", "_true")) %>%
  mutate(
    mean_diff = mean - mean_true,
    median_diff = median - median_true,
    upper_95_diff = upper_95 - upper_95_true,
    method = "Informed"
  )

df_diff_u <- df_tv_deconv_u %>%
  left_join(df_tv_true, by = "sample_size", suffix = c("", "_true")) %>%
  mutate(
    mean_diff = mean - mean_true,
    median_diff = median - median_true,
    upper_95_diff = upper_95 - upper_95_true,
    method = "Uninformed"
  )

df_diff_c <- df_tv_c %>%
  left_join(df_tv_true, by = "sample_size", suffix = c("", "_true")) %>%
  mutate(
    mean_diff = mean - mean_true,
    median_diff = median - median_true,
    upper_95_diff = upper_95 - upper_95_true,
    method = "convoluted"
  )

df_all <- rbind(df_diff_u,df_diff_i,df_diff_c)
df_all$method <- factor(df_all$method,
                        levels = c("convoluted", "Uninformed", "Informed"))

```
```{r}
# df_all <- rbind(
#   df_tv_c %>% mutate(method = "Convolved"),
#   df_tv_deconv_i %>% mutate(method = "Deconvolved informed"),
#   df_tv_deconv_u %>% mutate(method = "Deconvolved uninformed")
# )

p8 <- ggplot(df_all, aes(x = factor(sample_size), y = mean_diff, fill = method)) +
  geom_col(position = "dodge") +
  facet_wrap(~model) +
  labs(
    title = "Excess TV Distance Relative to the True Distribution sample",
    x = "Sample size",
    y = "Difference in Mean TV Distance",,
    fill = "method"
  ) +
  scale_fill_manual(values = c(
    "convoluted" = "black",
    "Informed" = "green",
    "Uninformed" = "darkgreen"
  )) +
  theme_minimal(base_size = 14)
p8
ggsave(
  filename = "../figures/barplot_c_u_i_TV_minus_baseline_p8.pdf",
  plot = p8,
  width = 6,
  height = 4
)
```

```{r}
# t-test
sim_results_c
deconvolved_samples_u
deconvolved_samples_i

compute_tv_all_per_replica <- function(sim_results, stable_dist) {
  
  lapply(seq_along(sim_results), function(i) {
    
    true_dist <- stable_dist
    
    lapply(sim_results[[i]], function(mat) {
      tv_vec <- apply(mat, 2, function(counts) {
    est_dist <- counts / sum(counts)
    tv_distance(est_dist, true_dist)
  })
    })
    
  })
}
```
```{r}
compute_tv_all_per_replica <- function(sim_results, stable_dist) {
  
  lapply(seq_along(sim_results), function(i) {
    
    true_dist <- stable_dist
    
    lapply(sim_results[[i]], function(mat) {
      tv_vec <- apply(mat, 2, function(counts) {
    est_dist <- counts / sum(counts)
    tv_distance(est_dist, true_dist)
  })
    })
    
  })
}
tv_per_rep_c <- compute_tv_all_per_replica(sim_results_c,stable_dist)
tv_per_rep_u <- compute_tv_all_per_replica(deconvolved_samples_u,stable_dist)
tv_per_rep_i <- compute_tv_all_per_replica(deconvolved_samples_i,stable_dist)

results_t_test <- data.frame()

for (i in seq_along(tv_per_rep_u)) {
  for (j in seq_along(tv_per_rep_u[[i]])) {

    c_vals <- tv_per_rep_c[[i]][[j]]
    u_vals <- tv_per_rep_u[[i]][[j]]
    i_vals <- tv_per_rep_i[[i]][[j]]

    test_i_u <- t.test(u_vals, i_vals, paired = TRUE)
    test_i_c <- t.test(c_vals, i_vals, paired = TRUE)
    test_u_c <- t.test(c_vals, u_vals, paired = TRUE)

    results <- rbind(results, data.frame(
      model = i,
      sample_size = j,

      mean_c = mean(c_vals),
      mean_u = mean(u_vals),
      mean_i = mean(i_vals),

      diff_i_u = mean(i_vals - u_vals),
      diff_i_c = mean(i_vals - c_vals),
      diff_u_c = mean(u_vals - c_vals),

      p_i_u = test_i_u$p.value,
      p_i_c = test_i_c$p.value,
      p_u_c = test_u_c$p.value
    ))
  }
}
results$model <- factor(results$model,
                     levels = 1:4,
                     labels = c("A", "B", "C", "D"))
  
results$sample_size <- factor(results$sample_size,
                           levels = 1:6,
                           labels = c("10", "25", "50", "100", "150", "200"))
  
# results
```

```{r}
results_long <- bind_rows(
  data.frame(
    model = results$model,
    sample_size = results$sample_size,
    comparison = "I-U",
    diff = round(results$diff_i_u,3),
    p_value = results$p_i_u
  ),
  data.frame(
    model = results$model,
    sample_size = results$sample_size,
    comparison = "I-C",
    diff = round(results$diff_i_c, 3),
    p_value = results$p_i_c
  ),
  data.frame(
    model = results$model,
    sample_size = results$sample_size,
    comparison = "U-C",
    diff = round(results$diff_u_c,3),
    p_value = results$p_u_c
  )
)
results_long <- results_long %>%
  mutate(
    stars = case_when(
      p_value < 0.001 ~ "***",
      p_value < 0.01  ~ "**",
      p_value < 0.05  ~ "*",
      TRUE ~ ""
    )
  )

top10 <- results_long %>%
  select(-p_value) %>%
  arrange(desc(abs(diff))) %>%
  slice(1:10)

# Export to ..\tables
table_tex <- kable(top10, format = "latex", booktabs = TRUE)

writeLines(table_tex, "../tables/diff_tv_top10_t4.tex")
```
### Population framskrivning
```{r}
#simulera från convolverad fel-modeller 
#simulation from true dist

# convolute 

convolved_sim <- lapply(missclass_all, function(K) {
  
  lapply(sim_results_true_dist, function(sim_list) {
    
    lapply(sim_list, function(mat) {
      
      out <- matrix(0, nrow = nrow(mat), ncol = ncol(mat))
      
      for (j in seq_len(ncol(mat))) {
        out[, j] <- simulate_convolved_counts(mat[, j], K)
      }
      
      out
    })
    
  })
  
})
# #  summary statistic of TV from sim_results
# tv_result_summary_deconv_i <-  compute_tv_all(deconvolved_samples_i, stable_dist)


#Simulera från s-paramter: 100 replikat med stickprovsstorlek n=(10,25,50,100,150,200) x_n-stickprov

#Framskriv 1-10 generationer L^i x_n = x_n^(i) i = 1,2,...,10

# ta medelvärde, median och tvåsidigt ci baserat på datat 
```

## EM-alg Skatta parameter p for a decon for a u-init

```{r}
deconvolved_p_samples_u <- Map(function(model_list, K) {
  
  lapply(model_list, function(scenario_list) {
    
    lapply(scenario_list, function(mat) {
      
      stopifnot(is.matrix(mat))  # IMPORTANT DEBUG SAFETY
      
      apply(mat, 2, function(col) {
        em_deconvolve(col, K = K)$p
      })
      
    })
    
  })
  
}, convolved_sim, missclass_all)
```

## EM-alg Skatta parameter p for a decon for informed init

```{r}
deconvolved_p_samples_i <- Map(function(model_list, K) {
  
  lapply(model_list, function(scenario_list) {
    
    lapply(scenario_list, function(mat) {
      
      stopifnot(is.matrix(mat))  # IMPORTANT DEBUG SAFETY
      
      apply(mat, 2, function(col) {
        em_deconvolve(col, K = K, init=TRUE)$p
      })
      
    })
    
  })
  
}, convolved_sim, missclass_all)
```

```{r}
# normalized convolved sim 

normalized_convolved_sim <- lapply(convolved_sim, function(model_list) {
  
  lapply(model_list, function(sample_list) {
    
    lapply(sample_list, function(mat) {
      
      stopifnot(is.matrix(mat))
      
      col_sums <- colSums(mat)
      
      # avoid division by zero
      col_sums[col_sums == 0] <- 1
      
      sweep(mat, 2, col_sums, `/`)
    })
    
  })
  
})
```

## forcasting 

```{r}

```

```{r}
#Plot part
leslie_p <- generate_leslie_powers(L, 100) 

f_proj_c <- compute_leslie_projection(deconvolved_p_samples_i[[1]][[1]][[1]], leslie_p, n_future_proj = 100)
summary_case <- summary_leslie_proj(f_proj_c)

ggplot(summary_case, aes(x = seq_len(nrow(summary_case)))) +
  
  geom_ribbon(aes(ymin = lower_95, ymax = upper_95),
              fill = "skyblue", alpha = 0.3) +
  
  geom_line(aes(y = mean), color = "blue", linewidth = 1) +
  
  geom_line(aes(y = median), color = "red", linetype = "dashed") +
  
  labs(x = "Time step", y = "Population") +
  theme_minimal()
```

